---
title: "Nanopore raw directory clean-up"
output: html_notebook
---



```{r}
library(tidyverse)
library(readxl)
```

```{r}
nano_dir <- read_excel("~/Desktop/20180212_nano_raw_cleanup_list.xlsx")
```

```{r}
dir_types <- nano_dir %>% 
    group_by(results, aln, fast5, fastq) %>% 
    summarise(count = n()) %>% 
    arrange(-count)
```

```{r}
dir_types
```
## Actions
- directories with fast5, fastq, and aln - make sure fastq and bam for each fast5
- fast5 only - look into why they were not processed

```{r}
dir_types %>% left_join(nano_dir) %>% filter(count == 13)
```

- aln and fastq only - make sure these are from collaborators
```{r}
dir_types %>% left_join(nano_dir) %>% filter(count == 6)
```

- No sub-directories/ empty sub-directories
```{r}
dir_types %>% left_join(nano_dir) %>% filter(count == 5)
```


- Misc.
```{r}
dir_types %>% left_join(nano_dir) %>% filter(count < 3)
```

## fast5 - fastq file checks

```{r}
fast5_df <- read_lines("raw_fast5_list.txt") %>% 
    keep(str_detect, "-rw") %>%
    discard(str_detect,"pipeline.log") %>% 
    str_sub( start = 27) %>%
    str_replace("(?<=M|G|K|[:digit:])\\s.*201[78]\\s","&") %>%
    str_remove("it\\s|t\\s") %>%
    str_replace("(?<=M|G|K|[:digit:])\\s.*[:digit:][:digit:]:[:digit:][:digit:]", "&") %>%
    as_tibble() %>%
    separate(value, into = c("fast5_size","fast5_name"), sep = "&") %>% 
    mutate(base_name = str_remove(fast5_name, ".tar"))
```


```{r}
fastq_df <- read_lines("raw_fastq_list.txt") %>% 
    keep(str_detect, "fastq.gz") %>%
    discard(str_detect, "txt|tsv|combined") %>% 
    str_sub( start = 27) %>%
    str_replace("(?<=M|G|K|[:digit:])\\s.*201[78]\\s","&") %>%
    str_remove("it\\s|t\\s") %>%
    str_replace("(?<=M|G|K|[:digit:])\\s.*[:digit:][:digit:]:[:digit:][:digit:]", "&") %>%
    as_tibble() %>%
    separate(value, into = c("fastq_size","fastq_name"), sep = "&") %>% 
    mutate(base_name = str_remove(fastq_name, ".fastq.gz")) 
```


Check for missing fastq files
```{r}
fastq_df %>% full_join(fast5_df) %>% filter(is.na(fastq_name)) %>% 
    select(base_name) %>% 
    mutate(base_name = str_remove(base_name, "_[:digit:]{1,2}_[:digit:]")) %>% 
    distinct()
```

## Runs without fast5s missing fastqs
- HalfPrecipxyz
- FAK27084_ - I believe this is a prometion run or from the new version of MinKNOW
- gm24385_181004DNA1F - known issue with tars with lowercase and capital F

Look into the following
- gm24385_180207DNA2a - Known issue typo in run name
- gm24385_180807DNA1b - need to check for basecalling errors
- gm24385_180807DNA1c - need to check for basecalling errors
- gm24385_180807DNA1a - need to check for basecalling errors
- gm24385_181001DNA1a - known issue not entered in database correctly


## Runs with small or missing fastqs
Most small fastqs are from small fast5s
Check gm24385_180807DNA1b, gm24385_181205DNA2b_0, and gm24385_181205DNA2b_1 fastq files
```{r}
fastq_df %>% full_join(fast5_df) %>% 
    filter(!str_detect(fastq_size, "K|M"))
```

